# 3: Logistic regression

## Part 1 

The existence of this chapter concludes the first part.

## Part 2

This is a joined dataset of math and Portuguese students, aiming to predict the school performance of students by many different variables. Variables are as described [here](https://archive.ics.uci.edu/ml/datasets/Student+Performance), but with the following changes:

Lists were joined by fixed value variables, and the rest variables were combined by averaging; for example, G1—G3 display the averages of 1st, 2nd and final period grades from both math and Portuguese classes. New columns alc_use (average values of students' weekday and weekend alcohol consumption values, range being 1—5 from low to very high) and high_use (displays if this average alcohol consumption value is higher than 2 or not) were created.

```{r}
# Reading the wrangled data excel
library(openxlsx); library(dplyr)
alc <- read.xlsx("~/R/IODS-project/data/pormath.xlsx")

# Sneak peek to the dimensions and column names of the wrangled dataset
dim(alc)
colnames(alc)

```

## Part 3

The purpose of this exercise is to study how the high/low alcohol consumption relates to a few other variables. I chose variables G3 (= average of the final grade in math and Portuguese courses), absences from school, health, and going out with friends. My hypothesis is that high alcohol consumption correlates *negatively* with higher grades and health, and *positively* with absences from school and going out with friends. I will also explore how alcohol consumption relates to reported study times.

## Part 4

Let's explore the chosen variables and their relationship with high/low alcohol consumption!

```{r}
g1 <- ggplot(alc, aes(x = high_use, y = G3, col = sex))
g1 + geom_boxplot() + ggtitle("Final grade (G3) vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Final grade")
```

My hypothesis seems to work in the case of this first variable; higher alcohol consumption correlates indeed with lower grades. However, the effect is much more pronounced in boys than it is in girls.

```{r}
g2 <- ggplot(alc, aes(x = high_use, y = absences, col = sex))
g2 + geom_boxplot() + ggtitle("Absences from school vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Absences")

g2_2 <- ggplot(data = alc, aes(x = absences, col=sex))
g2_2 + geom_bar()
```

Second variable supports as well the hypothesis, but the scale in this variable is so big that the few outliers with very high absence values shrink the distribution boxes visually. Bar plot was drawn to display distribution more clearly. Curiously, the alcohol consumption effect was again more pronounced in boys when compared to girls.

```{r}
g3 <- ggplot(alc, aes(x = high_use, y = health, col = sex))
g3 + geom_boxplot() + ggtitle("Health vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Health")
```

Third variable proved to be interesting... My hypothesis certainly was not correct on this one, as it seems that boys' health is not affected at all by high alcohol consumption in these data, and girls median health value is actually *higher* in the high alcohol consumption group, even if the lower quartile is a bit more spread out towards low health values at the same time.

```{r}
g4 <- ggplot(alc, aes(x = high_use, y = goout, col = sex))
g4 + geom_boxplot() + ggtitle("Going out with friends vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Going out with friends")
```

Fourth, and final, variable - going out with friends - definitely supports the hypothesis. Those students that consume more alcohol have a tendency to go out with their friends more as well.

Finally, I'll take a look to the relationship between weekly study time and alcohol consumption:
```{r}
alc %>% group_by(studytime) %>% summarise(alcohol_usage = mean(alc_use))
```
Clearly, there is a trend where higher study time correlates with lower amount of alcohol consumed. Maybe they don't have as much time to drink. :)

## Part 5



## Part 6
## Part 7
## Part 8