# 3: Logistic regression

## Part 1 

The existence of this chapter concludes the first part.

## Part 2

This is a joined dataset of math and Portuguese students, aiming to predict the school performance of students by many different variables. Variables are as described [here](https://archive.ics.uci.edu/ml/datasets/Student+Performance), but with the following changes:

Lists were joined by fixed value variables, and the rest variables were combined by averaging; for example, G1—G3 display the averages of 1st, 2nd and final period grades from both math and Portuguese classes. New columns alc_use (average values of students' weekday and weekend alcohol consumption values, range being 1—5 from low to very high) and high_use (displays if this average alcohol consumption value is higher than 2 or not) were created.

```{r}
# Reading the wrangled data excel
library(openxlsx)
alc <- read.xlsx("~/R/IODS-project/data/pormath.xlsx")

# Sneak peek to the dimensions and column names of the wrangled dataset
dim(alc)
colnames(alc)

```

## Part 3

The purpose of this exercise is to study how the high/low alcohol consumption relates to a few other variables. I chose variables G3 (= average of the final grade in math and Portuguese courses), absences from school, health, and going out with friends. My hypothesis is that high alcohol consumption correlates *negatively* with higher grades and health, and *positively* with absences from school and going out with friends. 

## Part 4

Let's explore the chosen variables and their relationship with high/low alcohol consumption!

```{r}
g1 <- ggplot(alc, aes(x = high_use, y = G3, col = sex))
g1 + geom_boxplot() + ggtitle("Final grade (G3) vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Final grade")
```
The first variable

```{r}
g2 <- ggplot(alc, aes(x = high_use, y = absences, col = sex))
g2 + geom_boxplot() + ggtitle("Absences from school vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Absences")
```
Second

```{r}
g3 <- ggplot(alc, aes(x = high_use, y = health, col = sex))
g3 + geom_boxplot() + ggtitle("Health vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Health")
```
Third

```{r}
g4 <- ggplot(alc, aes(x = high_use, y = goout, col = sex))
g4 + geom_boxplot() + ggtitle("Going out with friends vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Going out with friends")
```
Fourth, and final, variable

## Part 5
## Part 6
## Part 7
## Part 8