# RStudio Exercise 2

*Here I describe the work that I have done.*

## Part 1 
Here is the dataset that I wrangled in the previous part of the excercise. The full analysis dataset table can be read with read.table("data/learning2020.txt"), but that would be a ton of hard-to-read data to display in this diary page, so let's just take a closer look at the structure and dimensions of what we've got.

```{r}
# assigning the data to something easier to write in codes
lrn2020 <- read.table("data/learning2020.txt")

# dimensions of data: 183 observations and 7 variables
dim(lrn2020)

# structure of data: str() / gives you structural info and the amount of obs. & variables
str(lrn2020)

# structure of data: head() / I like this a bit more than str() for structural info, easier to read!
head(lrn2020)

```
## Part 2 
Show a graphical overview of the data and show summaries of the variables in the data. Describe and interpret the outputs, commenting on the distributions of the variables and the relationships between them. 
```{r}
# accessing GGally-library 
library(GGally)

# accessing ggplot2-library
library(ggplot2)

# creating plot matrix "p"
p <- ggpairs(lrn2020, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

#printing out the created matrix "p"
p

```

## Part 3
Choose three variables as explanatory variables and fit a regression model where exam points is the target (dependent) variable. Show a summary of the fitted model and comment and interpret the results. Explain and interpret the statistical test related to the model parameters. If an explanatory variable in your model does not have a statistically significant relationship with the target variable, remove the variable from the model and fit the model again without it. 
```{r}

# creating model with points versus attitude, strategic and deep learning
model_x <- lm(Points ~ Attitude + stra + deep, data = lrn2020)

summary(model_x)

```
## Part 4 
Using a summary of your fitted model, explain the relationship between the chosen explanatory variables and the target variable (interpret the model parameters). Explain and interpret the multiple R squared of the model.
```{r}

```
## Part 5 
Produce the following diagnostic plots: Residuals vs Fitted values, Normal QQ-plot and Residuals vs Leverage. Explain the assumptions of the model and interpret the validity of those assumptions based on the diagnostic plots.
```{r}
date()
```

