# 6: Analysis of longitudinal data

For the last exercise, we wrangled two datasets into long form from the original wide form, in order to allow us to perform analysis on them. 

Basically this means that we broke joined data in columns (we had certain time intervals as colums containing the corresponding measured results) to separate these time periods to their own column, resulting in four variables for each set; "treatment", "subject", "bprs" and "week" for BPRS dataset, describing response in measured BPRS test results to two different treatments to schizophrenia; and "ID", "Group", "Weight" and "Time" in RATS data which investigates measured rat weights after starting one of three possible diet groups.

As "week" and "time" are now their own variables, instead of "week0, week1, week2" etc., we can perform analysis that shows us measured responses over time. I ended up adding a new column, id, to the BPRS data, as this allowed me to plot different treatment groups simultaneously - the two treatment groups contained identical 1-20 subjects, whereas the new id category treats each individual response to a certain treatment as unique.

```{r, warning=F, message=F}
library(dplyr); library(ggplot2); library(lme4)

# Fetching the wrangled data
data_B <- read.table("data/bprs.txt")
data_R <- read.table("data/rats.txt")

# Assigning separate ID:s for same subjects but different treatments
# so that we have ID:s 1-20 for treatment 1 and 21-40 for treatment 2 
data_B$subject <- as.integer(data_B$subject)
data_B$id <- 
  ifelse(data_B$treatment==1,
         data_B$subject,data_B$subject+20)

# Factoring categorical variables again, as R interprets them as integers
data_B$treatment <- factor(data_B$treatment)
data_B$subject <- factor(data_B$subject)
data_R$ID <- factor(data_R$ID)
data_R$Group <- factor(data_R$Group)

# Glimpse of the data
glimpse(data_B)
glimpse(data_R)
summary(data_B)
summary(data_R)

```

**Part 1**

The assignment was to implement the analyses of Chapter 8 of MABS using the RATS data. 

```{r}

# Boxplot to visualize the raw data
ggplot(data_R, aes(x = as.factor(Time), y = Weight, fill = Group)) + 
  geom_boxplot()

# Plots of individual rat growth profiles within their corresponding groups

ggplot(data_R, aes(x = Time, y = Weight, group = ID, col = Group)) +
  geom_line(aes(linetype = Group)) + 
  facet_grid(. ~ Group, labeller = label_both) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "none")

# Standardizing weight

data_R <- data_R %>%
  group_by(Time) %>%
  mutate(Weight_std = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()

# Plotting again with standardized weights

ggplot(data_R, aes(x = Time, y = Weight_std, group = ID, col = Group)) +
  geom_line(aes(linetype = Group)) + 
  facet_grid(. ~ Group, labeller = label_both) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Standardized weight") +
  theme(legend.position = "none")

```
\
Text

$$se=\frac{sd(x)}{\sqrt{n}}$$

```{r}

# Summary data with mean and standard error of weight by group and time 
n <- data_R$Time %>% unique() %>% length()

rats_s <- data_R %>%
  group_by(Group, Time) %>%
  summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>%
  ungroup()

# Glimpse the data
glimpse(rats_s)

# Plot the mean profiles
ggplot(rats_s, aes(x = Time, y = mean, col = Group)) +
  geom_line(size=1) +
  geom_point(size=3, shape=10) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se)) +
  scale_y_continuous(name = "mean(weight) +/- se(weight)")

```

\
Text

```{r}

# Summary dataset by treatment and subject with mean as the summary variable (ignoring baseline Time = 1).
rats_sum <- data_R %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(rats_sum)

# Draw a boxplot of the mean versus treatment
ggplot(rats_sum, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "Mean weight after the start of diets")

```
\
Text

```{r}

# Creating a new dataset by filtering the three outliers 
rats_sum2 <- rats_sum %>%
  filter(ID!=2 & ID!=12 & ID!=13)  # Outlier ID's

ggplot(rats_sum2, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "Mean weight after the start of diets")  
  
```
\
Text

```{r, warning=F, message=F}
# Add the baseline from the original data as a new variable to the summary data
baselines <- 
  data_R %>% filter(Time==1) %>%
  rename(baseline=Weight)

rats_b <- rats_sum %>%
  mutate(baselines)

anova(lm(mean ~ baseline + mean, data = rats_b))

```



**Part 2**

Implement the analyses of Chapter 9 of MABS using the BPRS data. 

```{r}

# Plot of BPRS scores of subjects against time, colored by treatment group

ggplot(data_B, aes(x = week, y = bprs, group = id)) +
  geom_line(aes(col = treatment)) +
  scale_x_continuous(name = "Week") +
  scale_y_continuous(name = "BPRS score") +
  theme(legend.position = "top")

# Standardizing the variable "bprs" to "bprs_std", plotting again

data_B <- data_B %>%
  group_by(week) %>%
  mutate(bprs_std = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()

ggplot(data_B, aes(x = week, y = bprs_std, group = id)) +
  geom_line(aes(col = treatment, linetype = treatment)) +
  scale_y_continuous(name = "Standardized BPRS score")

```
\
Text

```{r, warning=F, message=F}
# Regression model
bprs_reg <- lm(bprs ~ week + treatment, 
               data = data_B)

summary(bprs_reg)

# Random intercept model 1
bprs_ref1 <- lmer(bprs ~ week + treatment + (1 | id), 
                 data = data_B, REML = FALSE)
summary(bprs_ref1)

# Random intercept model 2
bprs_ref2 <- lmer(bprs ~ week + treatment + (week | id), 
                 data = data_B, REML = FALSE)
summary(bprs_ref2)

# Random intercept model 2
bprs_ref3 <- lmer(bprs ~ week + treatment +treatment*week + (week | id), 
                 data = data_B, REML = FALSE)
summary(bprs_ref3)

# ANOVA test on the models
anova(bprs_ref1,bprs_ref2)
anova(bprs_ref2,bprs_ref3)

```
\
Text

```{r}
# Visualization of random intercept model "bprs_ref2"
data_B$fitted2 <- fitted(bprs_ref2)

ggplot(data_B, aes(x = week, y = fitted2, group = id)) +
  geom_line(aes(col = treatment, linetype = treatment)) +
  scale_y_continuous(name = "Fitted BPRS") +
  theme(legend.position = "top")

```
\
Text
