# 4: Clustering and classification
## Part 1

In this exercise we created ‘chapter4.Rmd’ to perform the clustering and classification analysis in.

## Part 2

Here I initialize the necessary libraries and load the analysis data from MASS library. The dataset consists of 14 variables with 506 observations about housing values in suburbs of Boston.

```{r}
# Libraries
library(MASS)
library(corrplot)
library(dplyr)
library(ggplot2)

# Load the Boston data from MASS:
data("Boston")

colnames(Boston)

```

The variables are **crim** (crime rate), **zn** (residential land), **indus** (non-retail business acres), **chas** (dummy variable), **nox** (nitrogen oxides concentration in ppm), **rm** (average rooms), **age** (built prior to 1940), **dis** (distances to employment centres), **rad** (accessibility to radial highways), **tax** (tax rate), **ptratio** (pupil-teacher ratio), **black** (relative proportion of blacks), **lstat** (lower status %) and **medv** (owner-occupied homes).

```{r}
# Exploring the data:
str(Boston)

```


## Part 3

```{r}

```

## Part 4

```{r}

```

## Part 5

```{r}

```

## Part 6

```{r}

```

## Part 7

```{r}

```


