# 3: Logistic regression

**Part 1** 

The existence of this chapter concludes the first part of the exercise.

**Part 2**

This is a joined dataset of math and Portuguese students, aiming to predict the school performance of students by many different variables. Variables are as described [here](https://archive.ics.uci.edu/ml/datasets/Student+Performance), but with the following changes:

Lists were joined by fixed value variables, and the rest variables were combined by averaging; for example, G1—G3 display the averages of 1st, 2nd and final period grades from both math and Portuguese classes. New columns alc_use (average values of students' weekday and weekend alcohol consumption values, range being 1—5 from low to very high) and high_use (displays if this average alcohol consumption value is higher than 2 or not) were created.

```{r}
# Reading the wrangled data excel
library(openxlsx); library(dplyr)
alc <- read.xlsx("~/R/IODS-project/data/pormath.xlsx")

# Sneak peek to the dimensions and column names of the wrangled dataset
dim(alc)
colnames(alc)

```

**Part 3**

The purpose of this exercise is to study how the high/low alcohol consumption relates to a few other variables. I chose variables G3 (= average of the final grade in math and Portuguese courses), absences from school, health, and going out with friends. My hypothesis is that high alcohol consumption correlates *negatively* with higher grades and health, and *positively* with absences from school and going out with friends. I will also explore how alcohol consumption relates to reported study times.

**Part 4**

Let's explore the chosen variables and their relationship with high/low alcohol consumption!

```{r}
g1 <- ggplot(alc, aes(x = high_use, y = G3, col = sex))
g1 + geom_boxplot() + ggtitle("Final grade (G3) vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Final grade")
```

My hypothesis seems to work in the case of this first variable; higher alcohol consumption correlates indeed with lower grades. However, the effect is much more pronounced in boys than it is in girls.

```{r}
g2 <- ggplot(alc, aes(x = high_use, y = absences, col = sex))
g2 + geom_boxplot() + ggtitle("Absences from school vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Absences")

g2_2 <- ggplot(data = alc, aes(x = absences, col=sex))
g2_2 + geom_bar()
```

Second variable supports as well the hypothesis, but the scale in this variable is so big that the few outliers with very high absence values shrink the distribution boxes visually. Bar plot was drawn to display distribution more clearly. Curiously, the alcohol consumption effect was again more pronounced in boys when compared to girls.

```{r}
g3 <- ggplot(alc, aes(x = high_use, y = health, col = sex))
g3 + geom_boxplot() + ggtitle("Health vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Health")
```

Third variable proved to be interesting... My hypothesis certainly was not correct on this one, as it seems that boys' health is not affected at all by high alcohol consumption in these data, and girls median health value is actually *higher* in the high alcohol consumption group, even if the lower quartile is a bit more spread out towards low health values at the same time.

```{r}
g4 <- ggplot(alc, aes(x = high_use, y = goout, col = sex))
g4 + geom_boxplot() + ggtitle("Going out with friends vs. alcohol consumption") + xlab("High use of alcohol") + ylab("Going out with friends")
```

Fourth, and final, variable - going out with friends - definitely supports the hypothesis. Those students that consume more alcohol have a tendency to go out with their friends more as well.

Finally, I'll take a look to the relationship between weekly study time and alcohol consumption:
```{r}
alc %>% group_by(studytime) %>% summarise(alcohol_usage = mean(alc_use))
```
Clearly, there is a trend where higher study time correlates with lower amount of alcohol consumed. Maybe they don't have as much time to drink. :)

**Part 5**

Here I fit a logistic regression model to the variables, first including all four from my hypothesis. I will also take a look to the odds ratios for all four of them. Finally, I fit a new model excluding the non-significant variables.

```{r}
model_alc1 <- glm(high_use ~ G3 + absences + health + goout, data = alc, family = "binomial")
summary(model_alc1)
coef(model_alc1)

```

From the P-values we can see that absences and going out with friends are statistically significant (P<0.05), and thus the odds ratios and confidence intervals should be significant. Grades (G3) and health are not significant, although health (P=0.09) is not very far. Coefficients (intercept) show the predicted mean value of y when x = 0. Let's see what the odds ratios and confidence intervals look like:

```{r}
# odds ratios (OR) & confidence intervals (CI)
OR <- coef(model_alc1) %>% exp
CI <- confint(model_alc1) %>% exp
# Print both
cbind(OR, CI)
```
From odds ratios can be interpreted that students that go out more often are 2.064 times more likely to be heavy drinkers.The more healthy and more absent (from school) ones seem to be just slightly more inclined to report high alcohol use values, and highly scored ones are a bit less (OR<1) likely to be heavy drinkers. The confidence intervals are significant if number 1 cannot be fitted between the 2.5 % and 97.5 % values; this is the case in absences and, especially, in "going out with friends" values.

Right. Let's see what our model looks like with the two significant variables:

```{r}
model_alc2 <- glm(high_use ~ absences + goout, data = alc, family = "binomial")
summary(model_alc2)
```
P values are very nice now, and Min-Max / 1Q-3Q are not too far off (each pair should be as far from 0, if the model fits well). 

**Part 6**

Here I test the predictive power of this model.

```{r}
# Explore the predictive power of the model: predict the probability of high_use:
probabilities <- predict(model_alc2, type = "response")

# Add the predicted probabilities to the dataset 'alc':
alc <- mutate(alc, probability = probabilities)

# Use the probabilities to make a prediction of high_use:
alc <- mutate(alc, prediction = probability > 0.5)

# Tabulate the target variable versus the predictions:
table(high_use = alc$high_use, prediction = alc$prediction) %>% addmargins()
```
Here we can see how 236 of 259 not-high alcohol users were predicted by the model. However, only 65 of 111 high users were predicted. Thus, this model seems more accurate in predicting which students are *not* likely to consume lots of alcohol on weekly basis, but misses some of the ones that actually do so.

```{r}
# plot of 'high_use' versus 'probability' in 'alc'
g <- ggplot(alc, aes(x = probability, y = high_use, col = prediction))
g + geom_point()
```
Here the same phenomena is illustrated graphically. There are less wrong predictions in the low ("false") alcohol consumption group than there are in the high usage group.

```{r}
# loss function (= mean prediction error)
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)}

# Average number of wrong predictions in the (training) data
loss_func(class = alc$high_use, prob = alc$probability)

# Comparing the result to a prediction that there would have been no high-users:
loss_func(class = alc$high_use, prob = 0)
loss_func(class = alc$high_use, prob = 1)
loss_func(class = alc$high_use, prob = 2)

```
Finally, the loss function shows us that this model missed ~23 % of the predictions, but simply guessing at would have missed 30 %. The guessing strategy's percentage goes rapidly down as prob values increase. Thus, this model is better than simply guessing in predicting high alcohol consumption in students.

**Part 7**

Simple 10-fold cross-validation

```{r}
# K-fold cross-validation
library(boot)
cv <- cv.glm(data = alc, cost = loss_func, glmfit = model_alc2, K = 10)

# average number of wrong predictions in the cross validation
cv$delta[1]
```
The error is 24 %.